\mytitle

### 2023-05-11 chapters/concl.tex
Fix bib/refs

There is still more work to be done in support of this thesis.
% TODO better transition sentence
%In addition to the cases presented, there are many more classes of systems
%that could benefit from program synthesis tools.
One interesting direction for work in this area is expanding on \jitsynth
to allow for some class of compiler optimizations.
In order to further improve the runtime performance of synthesized compilers,
\jitsynth could use superoptimization techniques to minimize the size
of discovered minicompilers.
Exploring how to best apply these techniques to improve \jitsynth's effectiveness
would be an interesting direction for future work.
%One interesting direction for work in this area is
%automating the implementation of blockchain staking protocols with program synthesis.
%Existing work has developed techniques for synthesizing more traditional distributed protocols, \todo{cite transit}
%and properties for staking protocols are specified in a fairly straightforward mannor,
%but complex required invariants on ___ make implementing staking mechanisms difficult.
%Automating this implementation step could prevent bugs in these protocols
%and the massive loss of funds that comes with these exploits.
Another promising direction lies in further improving \depsynth's usability.
Existing techniques for litmus test generation and partial interpretation of litmus tests
\todo{cite crashmonkey and memsynth}
could allow \depsynth to synthesize dependency rules more quickly.
% Distributed protocol synthesis: https://www.cis.upenn.edu/~alur/udupa-thesis.pdf
Finally, my overall thesis would be further supported by a larger
body of synthesis tools for a wide array of modern systems.
\cref{c:depsynth} discussed existing works in synthesis for
techniques for both memory models and distributed protocols,
but other promising system domains for program synthesis %application
include blockchain staking protocols and security monitors for IFC kernels.
%automating the implementation of blockchain staking protocols with program synthesis.
%Existing work has developed techniques for synthesizing more traditional distributed protocols, \todo{cite transit}
Properties for staking protocols are specified in a fairly straightforward mannor,
but complex invariants on the blockchain state make implementing staking mechanisms difficult.
\todo{cite https://academic.oup.com/rfs/article-abstract/34/3/1156/5868423, https://link.springer.com/chapter/10.1007/978-3-319-67816-0_17, https://link.springer.com/chapter/10.1007/978-3-319-63688-7_12}
Automating this implementation step could prevent bugs in staking protocols
and the massive loss of funds that comes with these exploits.
Information flow control (IFC) kernels give developers the ability to enforce
a variety of security properties by limiting the information flow between applications.
However, correctly implementing these restrictions is prohibitively difficult.
Existing work enables push-button verification of these security properties for IFC kernels,\todo{cite}
but still requires a manually-written implementation.
By synthesizing security monitors for IFC kernels that guarantee these properties,
the high implementation burden for IFC security could be removed entirely.
Overall, program synthesis tools can help systems developers build correct, performant software
more effectively and easily for a wide spectrum of system domains.

>>>>

There is still more work to be done in support of this thesis.
% TODO better transition sentence
%In addition to the cases presented, there are many more classes of systems
%that could benefit from program synthesis tools.
One interesting direction for work in this area is expanding on \jitsynth
to allow for some class of compiler optimizations.
In order to further improve the runtime performance of synthesized compilers,
\jitsynth could use superoptimization techniques to minimize the size
of discovered minicompilers.
Exploring how to best apply these techniques to improve \jitsynth's effectiveness
would be an interesting direction for future work.
%One interesting direction for work in this area is
%automating the implementation of blockchain staking protocols with program synthesis.
%Existing work has developed techniques for synthesizing more traditional distributed protocols, \todo{cite transit}
%and properties for staking protocols are specified in a fairly straightforward mannor,
%but complex required invariants on ___ make implementing staking mechanisms difficult.
%Automating this implementation step could prevent bugs in these protocols
%and the massive loss of funds that comes with these exploits.
Another promising direction lies in further improving \depsynth's usability.
Existing techniques for litmus test generation and partial interpretation of litmus tests
\todo{cite crashmonkey and memsynth}
could allow \depsynth to synthesize dependency rules more quickly.
% Distributed protocol synthesis: https://www.cis.upenn.edu/~alur/udupa-thesis.pdf
Finally, my overall thesis would be further supported by a larger
body of synthesis tools for a wide array of modern systems.
\cref{c:depsynth} discusses existing works in synthesis for
techniques for both memory models and distributed protocols,
but other promising system domains for program synthesis %application
include blockchain staking protocols and security monitors for IFC kernels.
%automating the implementation of blockchain staking protocols with program synthesis.
%Existing work has developed techniques for synthesizing more traditional distributed protocols, \todo{cite transit}
Security and liveness properties for staking protocols are specified in a fairly straightforward manner,
but complex invariants on the blockchain state make implementing staking mechanisms difficult.
\todo{cite https://academic.oup.com/rfs/article-abstract/34/3/1156/5868423, https://link.springer.com/chapter/10.1007/978-3-319-67816-0_17, https://link.springer.com/chapter/10.1007/978-3-319-63688-7_12}
Automating this implementation step could prevent bugs in staking protocols
and the massive loss of funds that comes with these exploits.
Information flow control (IFC) kernels give developers the ability to enforce
a variety of security properties by limiting the information flow between applications.
However, correctly implementing these restrictions is prohibitively difficult.
Existing work enables push-button verification of these security properties for IFC kernels,\todo{cite}
but still requires a manually-written implementation.
By synthesizing security monitors for IFC kernels that guarantee these properties,
the high implementation burden for IFC security could be removed entirely.
Overall, program synthesis tools can help systems developers build correct, performant software
more effectively and easily for a wide spectrum of system domains.

### 2023-05-11 chapters/intro.tex
Fix bib/refs

% Why is it hard?
Modern systems developers must tackle three challenges when constructing
systems that both meet high performance criteria and exhibit correct behavior.
First, developers must \textbf{specify} the intended behavior of their system at a high
level. This specification can be encoded in a variety of ways, from
English-text design documents to formal logic specifications.
Second, developers must \textbf{implement} their system in accordance with the defined
specificaiton. Generally, this implementation done at a lower level than the
specification and may contain details or optimizations not referenced by the specification.
Third, developers must \textbf{verify} that their implementation behaves correctly 
according to their specification. Much like the last two tasks, verification can be done
in one of several ways, from testing to manual or automated proofs to code review.
However, no matter the method for each of these three tasks,
all pose a significant burden of effort to systems developers.

@@

@@

% JitSynth: break up into minicompilers (TODO)
In order to synthesize such massive and complex systems,
\jitsynth considers the synthesis problem as one over per-instruction compilers
between \textit{abstract register machines} --- an abstraction over
in-kernel DSLs and other low-level languages introduced in \cref{c:jitsynth}.
\jitsynth decomposes the synthesis problem for per-instruction compilers
into multiple synthesis tasks for \textit{minicompilers}
that translate only one source instruction.% TODO include this line?
% JitSynth: design metasketch to describe common minicompiler patterns (TODO)
To prune and prioritize the search space for minicompilers,
\jitsynth takes advantage of a novel \textit{compiler metasketch}
that enables the efficient exploration of candidate minicompilers
using off-the-shelf satisfiability solvers.
With \jitsynth, we have synthesized three JIT compilers:
one from eBPF to RISC-V,
one from classic BPF to eBPF,
and one from libseccomp to eBPF.
Moreover, we have confirmed that the synthesized eBPF to RISC-V compiler
lacks bugs found in the existing Linux compiler.

\section{Automatic Crash Consistency with Synthesis}
Storage systems face an increasing demand for reliability.
From Linux file systems to Amazon cloud-based key-value stores,
storage systems must maintain the integrity of their data
even in the presence of system crashes.
This property of storage systems is known as \textit{crash consistency}.
Guaranteeing crash consistency is a highly non-trivial task for developers,
as it requires maintaining complex invariants that depend greatly on
storage system implementation details.
Moreover, failing to guarantee crash consistency may result in severe impacts,
including costly slowdowns and massive data losses. % TODO make more concrete
% TODO finish: Highlight what can go wrong if crash consistency is done incorrectly
% TODO
% DepSynth: consider storage mechanism separately from crash consistency mechanism (TODO)
% DepSynth: design rules language to describe any runtime dependencies (TODO)

@@

% TODO revise this paragraph
We take advantage of several novel abstractions in order enable
automatic crash consistency with \depsynth.
First, we assume a new \textit{angelic crash consistency} model
that delegates enforcement of write-before relationships to a
\textit{dependency-aware buffer cache}
that sits between the storage system software and hardware layers.
Second, we introduce a language for dependency rules
that allows for both \textit{expressive} and \textit{generalizable} rules.
Finally, we describe a new search for dependency rules
directed by input litmus tests.
These insights enable \depsynth to synthesize dependency rules for an existing
production key-value store at Amazon\todo{cite shardstore} in 49\todo{macro this} minutes.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Like software development for any domain, systems software development requires
three fundamental tasks:
\textbf{specifying} correct behavior at a high level,
\textbf{implementing} the system at a low level,
and \textbf{verifying} that the implementation correctly adheres to the specification.
Where systems software distinguishes itself is in the level of abstraction
at which the implementation operates.
Specifically, many systems must operate between layers of abstraction,
such as between source and target languages or between the application and
hardware layers of the operating system.
This means that manually implementing and verifying systems software requires
a high amount of effort from the developer, who must reason about the multiple abstraction
layers in addition to the execution of their implemented system.
Additionally, stringent requirements on performance for systems
further increase the difficulty of implementation and verification.

A well-studied technique for reducing the burden of both implementation
and verification from software designers is \textit{program synthesis}.
Program synthesis is an automated programming technique
that takes advantage of overlapping semantics between the three tasks described above
in order to automate software implementation and verification. 
To do so, synthesis tools take as input a specification and output both an implementation
as well as a certificate verifying that the implementation correctly behaves
according to the input specification.

The goal of my work is to automate system development with program synthesis.
Systems software is a particularly appealing target for program synthesis
due to the difficultly of manually implementing high-performance systems
that operate accross multiple abstraction layers.
However, for the same reasons that developers struggle to manually reason about systems,
creating tools to \textit{automatically} construct systems software poses
a uniquely difficult challenge.
Synthesis tools for systems must not only consider the space of candidate implementations,
but also the complex semantics of the abstractions that the system implementation operates between.
For JIT compiler synthesis tools, this means considering all possible executions of both source and target programs;
for crash consistency automation, this means reasoning about sets of disk writes
that a given storage system may issue,
as well as any possible reorderings of those writes made by the disk hardware.
Moreover, demands on the performance of systems further complicate this task,
restricting the space of acceptable outputs from a systems synthesis tool.

To overcome these unique challenges, this dissertation presents a variety of novel
\textit{system decomposition} and \textit{metaprogramming abstraction} techniques.
%In this work, I identify two key observations that enable the creation of synthesis tools
\textbf{Program decomposition} allows synthesis tools to generate smaller target programs.
By considering systems as a conjunction of smaller pieces,
synthesis tools can search for candidate programs that implement a single piece.
The resulting synthesized program can be combined with either user-written code or
other synthesized system pieces, effectively reducing the developer burden.
% JitSynth: break up into minicompilers (TODO)
% DepSynth: consider storage mechanism separately from crash consistency mechanism (TODO)
Second, system-specific \textbf{metaprogram abstractions} enable synthesis tools to reason 
about candidate programs that themselves take DSL programs as input.
By reasoning about these simpler abstractions rather than the complex systems directly,
synthesis tools are able to more efficiently search over candidate programs.
% JitSynth: design metasketch to describe common minicompiler patterns (TODO)
% DepSynth: design rules language to describe any runtime dependencies (TODO)
In the next two sections, I describe how these observations enabled the creation of
synthesis tools for both in-kernel JIT compliers and for crash consistent storage systems.

My thesis is that \textit{program
synthesis is an effective tool for designing systems that are both correct and
performant}.  This dissertation explores this thesis over two particular types
of systems --- just-in-time compilers and crash consistent storage systems ---
and demonstrates how new synthesis tools can reduce implementation and verification
effort for developers despite the size and complexity of these systems.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if 0
% Why is it hard?
Modern systems developers must tackle three challenges when constructing
systems that both meet high performance criteria and exhibit correct behavior.
First, developers must \textbf{specify} the intended behavior of their system at a high
level. This specification can be encoded in a variety of ways, from
English-text design documents to formal logic specifications.
Second, developers must \textbf{implement} their system in accordance with the defined
specificaiton. Generally, this implementation done at a lower level than the
specification and may contain details or optimizations not referenced by the specification.
Third, developers must \textbf{verify} that their implementation behaves correctly 
according to their specification. Much like the last two tasks, verification can be done
in one of several ways, from testing to manual or automated proofs to code review.
However, no matter the method for each of these three tasks,
all pose a significant burden of effort to systems developers.

@@

\fi

@@

% JitSynth: break up into minicompilers (TODO)
In order to synthesize such massive and complex systems,
\jitsynth considers the synthesis problem as one over per-instruction compilers
between \textit{abstract register machines} --- an abstraction over
in-kernel DSLs and other low-level languages introduced in \cref{c:jitsynth}.
\jitsynth decomposes the synthesis problem for per-instruction compilers
into multiple synthesis tasks for \textit{minicompilers}
that translate only one source instruction. % TODO include this line?
% JitSynth: design metasketch to describe common minicompiler patterns (TODO)
To prune and prioritize the search space for minicompilers,
\jitsynth takes advantage of a novel \textit{compiler metasketch}
that enables the efficient exploration of candidate minicompilers
using off-the-shelf satisfiability solvers.
With \jitsynth, we have synthesized three JIT compilers:
one from eBPF to RISC-V,
one from classic BPF to eBPF,
and one from libseccomp to eBPF.
Moreover, we have confirmed that the synthesized eBPF to RISC-V compiler
lacks bugs previously found in the existing Linux compiler.

\section{Automatic Crash Consistency with Synthesis}
Storage systems face an increasing demand for reliability.
From Linux file systems to large cloud-based key-value stores,
storage systems must maintain the integrity of their data
even in the presence of system crashes.
This property of storage systems is known as \textit{crash consistency}.
Guaranteeing crash consistency is a highly non-trivial task for developers,
as it requires maintaining complex invariants that depend greatly on
storage system implementation details.
Moreover, failing to guarantee crash consistency may result in severe impacts,
including costly slowdowns and massive data losses. % TODO make more concrete
% TODO finish: Highlight what can go wrong if crash consistency is done incorrectly
% TODO
% DepSynth: consider storage mechanism separately from crash consistency mechanism (TODO)
% DepSynth: design rules language to describe any runtime dependencies (TODO)

@@

% TODO revise this paragraph
We take advantage of several novel abstractions in order enable
automatic crash consistency with \depsynth.
First, we assume a new \textit{angelic crash consistency} model
that delegates enforcement of write-before relationships to a
\textit{dependency-aware buffer cache},
which sits between the storage system software and disk hardware layers.
Second, we introduce a language for dependency rules
that allows for both \textit{expressive} and \textit{generalizable} rules.
Finally, we describe a new search for dependency rules
directed by input litmus tests.
These insights enable \depsynth to synthesize dependency rules for an existing
production key-value store at Amazon\todo{cite shardstore} in 49\todo{macro this} minutes.
